Accounting for uncertainty in simulating greenhouse gas emissions from dairy cattle in Kenya

### 2022-06-20 Manuscript_figures.Rmd
first (almost) complete draft

#install and load packages
# devtools::install_github("CWWhitney/uncertainty")
library(DiagrammeR)
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
library(uncertainty)
library(chillR)
library(raster)
library(reshape2)
library(patchwork)

knitr::write_bib(c(.packages(), 
                   'DiagrammeR',
                   'decisionSupport',
                   'ggplot2',
                   'kableExtra',
                   'magrittr',
                   'plyr',
                   'rmarkdown',
                   'uncertainty',
                   'chillR',
                   'raster',
                   'reshape2',
                   'patchwork'), 'references/packages.bib')
#resolves an issue with the scientific style of numbers,answer from @Paul Hiemstra: https://stackoverflow.com/a/25947542/4061993 
#from the documentation of ?options:
options(scipen=999)
```

```{r, warning=FALSE, eval=FALSE}
devtools::install_github("CWWhitney/uncertainty")
library(DiagrammeR)
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
library(uncertainty)
library(chillR)
library(raster)
library(reshape2)
library(patchwork)
```

@@

This equation takes a wide range of inputs. Many of these are usually not known precisely, yet the classic IPCC equation does not account for potential errors and uncertainties. Here, we express this uncertainty by defining distributions for each variable. The following table (**data/livestock_ghg_input_table.csv**) shows all the variables, as well as examples of distributions that may describe their quantities. Distributions are defined according to the distribution shape, as well as upper and lower bounds of 90% confidence intervals. We use the following distribution shapes:

@@

@@

Here's an example `scenarios` file:

```{r}
read.table("data/scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)   %>%
    kableExtra::kbl(digits = 2)  %>%
    kableExtra::kable_classic_2(full_width = F)

The first row ("Runs" in the first column) indicates the number of runs. Each subsequent row indicates which value from the input table should be modified, so the string in the first column has to correspond to a variable name in the input table. The second column indicates which values should be changed, with options being "lower", "upper", "distribution" and "both". If "both" is selected, both the lower and upper bounds are modified (this should only be done for variables with a constant distribution, since the replacement will be for both the upper and lower bound).

@@

read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)[1:20,1:5]  %>%
    kableExtra::kbl(digits = 2)  %>%
    kableExtra::kable_classic_2(full_width = F)

@@

@@

@@

kable(emissions_summary[,c(4,1,2,3)], booktabs = TRUE,
      col.names = c("Distribution attribute",
                    paste0("CO","$_2$","-eq farm","$^{-1}$"),
                    paste0("CO","$_2$","-eq head","$^{-1}$"),
                    paste0("CO","$_2$","-eq kg","$^{-1}$")),
      caption = "<p style='color:black'><i><b>Table S1.</b> Summary statistics of simulated greenhouse gas emissions from cattle across a population of cattle-keeping smallholder farms in Kenya</i></p>") %>% pack_rows(
  index = c("Emissions per farm" = 5, "Emissions per dairy-focused farm" = 5), background="lightgray") %>%
column_spec(2:4, width = "5cm") %>% row_spec(0, align = "c") %>% 
  add_header_above(c(" " = 1, "Emissions per farm" = 1, "Emissions per head" = 1, "Emissions intensity" = 1))

@@

```{r, fig.dim = c(8, 4), fig.cap="***Figure S2.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-head milk production levels (see caption to Fig. 5 for more details). The red line indicates a confidence threshold that is based on the 95% confidence interval of the loess regression line shown in Fig. 3a. Using this threshold for predicting changes in emissions intensity based on milk yield data would likely lead to a lot of inaccurate classifications.*"}

@@

```
```{r}
kable(breed_milk_summary, col.names = c("Breed", "Number of dairy cows", "Daily milk yield (kg)", "Intervention factor"),
      caption = "<p style='color:black'><i><b>Table S2.</b> Mean milk yield across all dairy cow breeds encountered during a farm survey in Kenya</i></p>") %>%
column_spec(2:4, width = "5cm")

@@

kable(intervention_summary, booktabs = TRUE,digits=0,
     # col.names = c("Intervention",
       #             paste0("CO","$_2$","-eq farm","$^{-1}$"),
        #            paste0("CO","$_2$","-eq head","$^{-1}$"),
         #           paste0("CO","$_2$","-eq kg","$^{-1}$")),
      caption = "<p style='color:black'><i><b>Table S3.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya. Impacts are evaluated against a true counterfactual baseline, in which all farms are identical except for the simulated impacts of the interventions.  </i></p>") %>%
  pack_rows(index = c("Impact on milk yield per head (%)" = 5, "Impact on emissions intensity (%)" = 5), background="lightgray") %>%
column_spec(2:6, width = "3cm")

@@

kable(intervention_end_summary, booktabs = TRUE,digits=0,
      caption = "<p style='color:black'><i><b>Table S4.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya. Results are based on comparison of a baseline scenario, where no farm was assumed to have implemented the interventions and an endline scenarios, where some farms adopted the mitigation measures. At both baseline and endline, variation in outcomes expected to result from random effects was simulated using Monte Carlo simulation (10 runs per farm, for 414 farms).</i></p>") %>%
  pack_rows(index = c("Impact on milk yield per head (%)" = 6, "Impact on emissions intensity (%)" = 6), background="lightgray") %>%
column_spec(2:6, width = "3cm")

>>>>

#install and load packages
# devtools::install_github("CWWhitney/uncertainty")
library(DiagrammeR)
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
library(uncertainty)
library(chillR)
library(raster)
library(reshape2)
library(patchwork)
library(DT)

knitr::write_bib(c(.packages(), 
                   'DiagrammeR',
                   'decisionSupport',
                   'ggplot2',
                   'kableExtra',
                   'magrittr',
                   'plyr',
                   'rmarkdown',
                   'uncertainty',
                   'chillR',
                   'raster',
                   'reshape2',
                   'patchwork'), 'references/packages.bib')
#resolves an issue with the scientific style of numbers,answer from @Paul Hiemstra: https://stackoverflow.com/a/25947542/4061993 
#from the documentation of ?options:
options(scipen=999)
options(DT.options = list(scrollY="300px",scrollX="300px", pageLength = 100, autoWidth = TRUE))

```{r, warning=FALSE, eval=FALSE}
devtools::install_github("CWWhitney/uncertainty")
library(DiagrammeR)
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
library(uncertainty)
library(chillR)
library(raster)
library(reshape2)
library(patchwork)
library(DT)
```

@@

This equation takes a wide range of inputs. Many of these are usually not known precisely, yet the classic IPCC equation does not account for potential errors and uncertainties. Here, we express this uncertainty by defining distributions for each variable. The following table (Table S1) shows all the variables, as well as examples of distributions that may describe their quantities. Distributions are defined according to the distribution shape, as well as upper and lower bounds of 90% confidence intervals. We use the following distribution shapes:

@@

***Table S1*** *Model input parameters for the Monte Carlo simulation, including the variable names used in the greenhouse gas emissions equation, a descriptive label, lower and upper bounds of 90% confidence intervals for the values, and the shape of the probability distribution that is used to describe the data. The IPCC_default column shows values used for producing (spuriously) the precise scenario. Note that some values are overwritten when producing farm-scale scenarios. Data can be downloaded in csv or Excel format or copied to the clipboard by pressing the respective buttons.*

input_parameters<-read.csv("data/livestock_ghg_input_table.csv")[,c(7,2,3,5,6,8)]
input_parameters<-input_parameters[which(!is.na(input_parameters$lower)),]

input_parameters %>%
  datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All")),
                           list(scrollX = "300px"),
                           autoWidth = TRUE),
            rownames=FALSE)
```

@@

Each row in this table indicates which value from the input table should be modified, so the string in the first column has to correspond to a variable name in the input table (additionally, the number of runs per scenario can be specified by a row with 'Runs' in the 'Variable' column, but this was not done here). The second column indicates which values should be changed, with options being "lower", "upper", "distribution" and "both". If "both" is selected, both the lower and upper bounds are modified (this should only be done for variables with a constant distribution, since the replacement will be for both the upper and lower bound).

@@

@@

***Table S2*** *Model input parameters for producing farm-scale scenarios for the Monte Carlo simulation, listing all variables that need to be updated. Data can be downloaded in csv or Excel format or copied to the clipboard by pressing the respective buttons.*

scenario_parameters<-read.csv("data/farm_scenarios.csv")

scenario_parameters %>%
  datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All")),
                           list(scrollX = "300px")),
            rownames=FALSE)
```

@@

The basic parameters for the simulations are shown in Table S1. Note that some of these parameters are overwritten when producing farm-scale scenarios.

Table S2 contains the scenario data derived from results of the farm survey. In running the Monte Carlo with the `scenario_mc` function, the initial values for all variables listed in this table (Table S1) are overwritten by those shown here (in Table S2) to generate farm-specific scenarios.

@@

kable(emissions_summary[,c(4,1,2,3)], booktabs = TRUE,
      col.names = c("Distribution attribute",
                    paste0("CO","$_2$","-eq farm","$^{-1}$"),
                    paste0("CO","$_2$","-eq head","$^{-1}$"),
                    paste0("CO","$_2$","-eq kg","$^{-1}$")),
      caption = "<p style='color:black'><i><b>Table S3.</b> Summary statistics of simulated greenhouse gas emissions from cattle across a population of cattle-keeping smallholder farms in Kenya</i></p>") %>% pack_rows(
  index = c("Emissions per farm" = 5, "Emissions per dairy-focused farm" = 5), background="lightgray") %>%
column_spec(2:4, width = "5cm") %>% row_spec(0, align = "c") %>% 
  add_header_above(c(" " = 1, "Emissions per farm" = 1, "Emissions per head" = 1, "Emissions intensity" = 1))

@@

```{r, warning=FALSE, fig.dim = c(8, 4), fig.cap="***Figure S2.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-head milk production levels (see caption to Fig. 5 for more details). The red line indicates a confidence threshold that is based on the 95% confidence interval of the LOESS regression line shown in Fig. 3a. Using this threshold for predicting changes in emissions intensity based on milk yield data would likely lead to a lot of inaccurate classifications.*"}

@@

```
```{r}
kable(breed_milk_summary, col.names = c("Breed", "Number of dairy cows", "Daily milk yield (kg)", "Intervention factor"),
      caption = "<p style='color:black'><i><b>Table S4.</b> Mean milk yield across all dairy cow breeds encountered during a farm survey in Kenya</i></p>") %>%
column_spec(2:4, width = "5cm")

@@

kable(intervention_summary, booktabs = TRUE,digits=0,
     # col.names = c("Intervention",
       #             paste0("CO","$_2$","-eq farm","$^{-1}$"),
        #            paste0("CO","$_2$","-eq head","$^{-1}$"),
         #           paste0("CO","$_2$","-eq kg","$^{-1}$")),
      caption = "<p style='color:black'><i><b>Table S5.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya. Impacts are evaluated against a true counterfactual baseline, in which all farms are identical except for the simulated impacts of the interventions.  </i></p>") %>%
  pack_rows(index = c("Impact on milk yield per head (%)" = 5, "Impact on emissions intensity (%)" = 5), background="lightgray") %>%
column_spec(2:6, width = "3cm")

@@

kable(intervention_end_summary, booktabs = TRUE,digits=0,
      caption = "<p style='color:black'><i><b>Table S6.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya. Results are based on comparison of a baseline scenario, where no farm was assumed to have implemented the interventions and an endline scenarios, where some farms adopted the mitigation measures. At both baseline and endline, variation in outcomes expected to result from random effects was simulated using Monte Carlo simulation (10 runs per farm, for 414 farms).</i></p>") %>%
  pack_rows(index = c("Impact on milk yield per head (%)" = 6, "Impact on emissions intensity (%)" = 6), background="lightgray") %>%
column_spec(2:6, width = "3cm")
